---
title: "1015_exercise"
author: "Chang"
date: "2021/10/15"
output: 
  html_document:
    toc: true
    toc_float: true
---
### exercise 0
Calculate BMI
```{r}
f <- function(w, h){
  BMI <- round(w/(h/100)^2, 2)
  print(
    paste(w, "(weight)/", round(h/100, 2), "^2 (square meter of height)=", BMI, sep = "")
  )
}
w <- sample(40:200, 1)
h <- sample(110:190, 1)
f(w, h)
```


### exercise 1 
寫一個程式，資料輸入後，會計算數列x的個數，平均及標準差。
```{r}
g <- function(x){
  seq_num <- length(x)
  mean <- round(mean(x), 2)
  sd <- round(sd(x), 2)
  return(list(seq_num, mean, sd))
}
length <- sample(1:50, 1)
x <- sample(1:100, length)
g(x)
```

### exercise 2
Write a funtion for Two sample t-test assuming that unknown population variances are equal. Return means, pooled variance, and t test statistic.
```{r}
ttest <- function(var1, var2){
  n1 <- length(var1)
  n2 <- length(var2)
  v1 <- var(var1)
  v2 <- var(var2)
  m1 <- mean(var1)
  m2 <- mean(var2)
  sp <- ((n1 - 1)*v1 + (n2 - 1)*v2)/(n1 + n2 - 2)
  t <- (m1 - m2)/sqrt(sp*(1/n1 + 1/n2))
  return(
       c(mean = c(m1,m2), 
         pooled.var = sp, 
         t.stat = t,
         # pt:probability test
         pvalue = 2*pt(-abs(t),df=n1+n2-2))
    )
}
var1 <- rnorm(10,3,1)
var2 <- rnorm(10,0,1)
ttest(var1, var2)
t.test(var1, var2, var.equal = T)
```

### exercise 3
請試寫R程式:抽撲克牌，算出的機率是否符合理論。
Draw a five-card poker hand from a standard deck of 52 cards, which contains 13 hearts(h), 13 diamonds(d), 13 clubs(c), and 13 spades(s). What is the probability that you get the same number of hearts as diamonds?
```{r}
# 五張牌要有相同數量的紅心和菱形牌，可能的情形有三種，各零、一、二張
# theoretical computation
numerator <- choose(26, 5) + choose(13, 1)*choose(13, 1)*choose(26, 3) + choose(13, 2)*choose(13, 2)*choose(26, 1)
denominator <- choose(52, 5)
numerator/denominator
```

```{r}
# one time
deck <- c(rep("h", 13), rep("d", 13), rep("c", 13), rep("s", 13))
result <- sample(deck, 5)
hearts <- sum(result == "h")
diamonds <- sum(result == "d")
success <- diamonds == hearts
data.frame(t(result), diamonds, hearts, success)

# several times
one <- function(){
  deck <- c(rep("h", 13), rep("d", 13), rep("c", 13), rep("s", 13))
result <- sample(deck, 5)
hearts <- sum(result == "h")
diamonds <- sum(result == "d")
success <- diamonds == hearts
data.frame(t(result), diamonds, hearts, success)
}

many <- function(NNN) {
  replicate(NNN, one()$success)
}

# calculate the probability
table(many(1000))
res <- as.data.frame(many(1000))
sum(res)/1000
```

```{r}
#check results(show the details of 10 experiments to make sure one is behaving correctly.)
do.call(rbind, lapply(1:10, function(nnn) {
  one()
})) ##do.call(fun,data)



```


### exercise 4
將年齡資料轉換為年齡群組1~20, 21~40, 41~60, 61歲以上，並編碼為A, B, C, D 。
```{r}
age <- sample(1:100, 20)
code <- ifelse(age >= 1 & age <= 20, "A", 
       ifelse(age >= 21 & age <= 40, "B",
              ifelse(age >= 41 & age <= 60, "C","D")))
head(cbind(age, code))
```

### exercise 5
計算n!
```{r}
# method1: using function()
f <- function(n){
  prod(seq(1,n))
}
n <- sample(1:10, 1)
cat(n,"! = ", f(n),"")

# method2: using function() and for loop
k <- 1
prodfcn <- function(n){
  for(i in 1:n){
  k <- k*i
  }
  return(k)
}
n <- sample(1:10, 1)
cat(n,"! = ", prodfcn(n),"")

# method3-1: using while
n <- sample(1:10, 1)
i <- 1
k <- 1
while(i <= n){
  k <- k*i
  i <- i + 1
}
cat(n,"! = ", k, "")

# method3-2: using while
f.while <- function(n){
  f <- 1
  t <- n
  while(t > 1){
    f <- f*t
    t <- t-1
  }
  return(f)
}
n <- sample(1:10, 1)
cat(n,"! = ", f.while(n),"")

# method4-1: using repeat
n <- sample(1:10, 1)
i <- 1
k <- 1
repeat{
  
  if (i > n) break
  k <- k*i
  i <- i + 1
}
cat(n,"! = ", k, "")

# method4-2: using repeat
f.repeat <- function(n){
  f <- 1
  t <- n
  repeat{
    if(t < 2) break
    f <- f*t
    t <- t-1
  }
  return(f)
}
n <- sample(1:10, 1)
cat(n,"! = ", f.repeat(n),"")

```

### exercise 6
投擲一枚公正的銅板，想知道總共需要投擲幾次才會出現三個正面。
```{r}
result <- c()
coin <- c(0,1)
i <- 1
while(sum(result) < 3){
  result[i] <- sample(coin, 1)
  i <- i + 1
}
length(result)
result
```
### exercise 7
計算Iris dataset data(iris) 的歐式距離(Euclidean Norm)。

```{r}
# R內建的距離函數
# dist(data, diag = T, upper=T)

n <- nrow(iris)
D <- matrix(0,n,n) #create a matrix to save the results
data <-iris[,-5]

for( i in 1:n){
  for(j in 1:n){
  D[i,j]<- sqrt(sum((data[i,]-data[j,])^2))
  }
}
D[3,3]
D[1,5]

```


### exercise 8
用R for loop 計算矩陣乘法
```{r}
A <- matrix(sample(1:5, 12, replace = T), 3, 4)
B <- matrix(sample(1:5, 12, replace = T), 4, 3)
C <-  matrix(0 , dim(A)[1] , dim(B)[2] , TRUE )
mat_mult <- function(A, B){
  for( i in 1:3){
    for( j in 1:3){
      C[i,j] <- sum(A[i,]*B[,j])
    }

  }
  return(C)
}
mat_mult(A, B)


# check
identical(A %*% B, mat_mult(A, B))
```



