---
title: "1015 R_function"
author: "Chang"
date: "2021/10/15"
output: 
  html_document:
    toc: true
    toc_float: true
---

### function
```{r}
df <- data.frame(
  # Generate 10 standard normal random numbers
  a <- rnorm(10),
  b <-  rnorm(10),
  c <-  rnorm(10),
  d <-  rnorm(10)
)

# basic method:
fcn <- function(x){
  (x - min(x))/(max(x) - min(x))
}
fcn(df$a)

# another method: use range()
# output of range: range[1] is min; range[2] is max
fcn1 <- function(x){
  minmax <- range(x, na.rm = T)
  ans <- (x - minmax[1])/(minmax[2] - minmax[1])
  return(ans)
}
fcn1(df$a)
```

### operator
```{r}
set.seed(1)
x <- runif(10, -1, 1)

# &：每一個值個別比較
0 <= x & x <= 0.5

# &&：只檢查第一個值
0 <= x && x <= 0.5 

# |：每一個值個別比較
0.4 >= x | x <= -0.2

# ||：只檢查第一個值
0.4 >= x || x <= -0.2

#滿足條件的換成NA
x[0 <= x & x <= 0.5] <- NA
x
```

### if_else

    Command:

    1.  if ('條件一'){
         "結果一"
        } else if ('條件二'){
         "結果二"
        } else{
         "結果三"
        }

    2.  ifelse(條件, "結果一", "結果二")

    3.  if ('條件') print("結果")

```{r}
# method 1:
f <- function(w, h){
  BMI <- round(w/(h/100)^2, 2)
  print(
    paste(w, "(weight)/", round(h/100, 2), "^2 (square meter of height)=", BMI, sep = "")
  )
  if (BMI < 18.5){
    print("underweight")
  } else if (BMI >= 18.5 & BMI < 24){
    print("normal")
  } else{
    print("overweight")
  }
}
w <- sample(40:200, 1)
h <- sample(110:190, 1)
f(w, h)

# method 2:
f0 <- function(w, h){
  BMI <- round(w/(h/100)^2, 2)
  print(
    paste(w, "(weight)/", round(h/100, 2), "^2 (square meter of height)=", BMI, sep = "")
  )
  ifelse(BMI < 18.5, "underweight", 
         ifelse(BMI >= 18.5 & BMI < 24, "normal", "overweight"))
}
f0(w, h)
```

### switch
可以同時輸入許多獨立的工作項目，再從中指定要執行哪一項。
```{r}
schedule <- function(x){
  switch (x,
  morning = "reading",
  noon = "nap",
  afternoon = "writing",
  night = "sleep")
}
schedule("noon")
```

### loops

    \t :空白格
    
    \n :換行
    
    \r :到下一行最開始處(少用)
    
#### for 
```{r}
# 從1加到100
result <- 0
for(i in 1:100){
  result <- result + i
}
result

# 九九乘法表
for (i in 1:9){
  for (j in 1:9){
    prod <- i*j
    cat(i,"x",j,"=",prod , "\t")
  }
  cat("\n")
}

```

#### while
符合條件會執行動作，直到條件不滿足為止。which()括號內放邏輯判斷句。
```{r}
# 從1加到100
i <- 1
result <- 0
while(i < 101){
  result <- result + i
  i <- i + 1
}
result


# 九九乘法表
i <-  1
while(i < 10){
  j <-  1
  while(j < 10){
    prod <- i*j
    cat(i,"x",j,"=",prod ,"\t")
    j <-  j + 1
  }
  cat("\n")
  i <-  i + 1
}
```

#### repeat and break
if (條件) break: 達到條件時，跳出迴圈。

```{r}
# 從1加到100
i <- 1
result <- 0
repeat{
  result <- result + i
  i <- i + 1
  
  # repeat 的條件可以寫在大括弧中的任何地方
  if (i > 100) break
  
}
result

# 九九乘法表
i <- 1
repeat{
  j <- 1
  repeat{
    
    if (j > 9) break
    
    prod <- i*j
    cat(i,"x",j,"=",prod ,"\t")
    j <-  j + 1
  }
  i <-  i + 1
  if (i > 9) break
  cat("\n")
}

```

#### next

* The next iteration of the loop is then executed.
    
* if (條件) next: 達到條件即時，跳過不執行該條件下的工作。
```{r}
for(i in c(1:6)){
  
  if(i == 3) next  #當i等於3的時候，省略此次迴圈(skip)的動作，從下一個i=4開始
  
  print(i)  
}
```
### apply 家族函數
https://chang-web.github.io/R-selflearning/R%20intermediate/R_intermediate.html#4

**apply(array, margin, ...)**
```{r}
mat <- array(1:20, c(4, 5)) 
mat

# count mean in columns: use apply()
apply(mat, 2, mean)

# count mean in columns: use for loop
X <- c()
for(i in 1:5){
  X[i] <- mean(mat[ ,i])
}
X

# count mean in columns: use function and sapply
M <- function(i){
  Mean <- mean(mat[ ,i])
  return(Mean)
}
sapply(1:5, M)

# use apply to count margin in rows
apply(mat, 1, sum)

# 找出每一行被3整除的數字
# 簡化版: apply(mat, 2,function(x) x[x %% 3 == 0])
apply(mat, 1, 
      function(x){
        x[x %% 3 == 0]
      }
)
```

```{r}
fcn <- function(x){ 
     return(sqrt(x)*10)  
 }
 
# 對每行元素逐一計算，依序擺入行中
apply(mat, 2, fcn)

# 對每行元素逐一計算，依序擺入行中(相當於前者轉置)
apply(mat, 1, fcn)

```
**lapply(list, function)**

* returns a list of the same length as X, each element of which is the result of applying FUN to the corresponding element of X.

**sapply(list, function)**

* sapply is a user-friendly version and wrapper of lapply by default returning a vector, matrix or an array if appropriate.

**vapply(list, function, FUN.VALUE = type, ...)**

* vapply is similar to sapply, but has a pre-specified type of return value, so it can be safer (and sometimes faster) to use.

* sapply(*, simplify = FALSE, USE.NAMES = FALSE) is equivalent to lapply(*) 

**tapply(vector, index, function)**
INDEX:The elements are coerced to factors by as.factor.
```{r}
library(tidyverse)
library(utils)
salesdata <- read.csv(file = 'D:/Chang-web/salesdata.csv')
# object:依照Product和Store分類，填入總消費平均。
res <- salesdata %>% 
  mutate(totalspend = Quantity*UnitPrice) %>%
  select(Product, Store, totalspend) 

# use tapply
product <- as.factor(res$Product)
store <- as.factor(res$Store)
spend <- as.numeric(res$totalspend)
object <- as.data.frame(
  round(tapply(spend, list(store, product), mean), 2)
)
object
```

```{r}
# review: use tidyverse
res %>% 
  group_by(Product, Store) %>% 
  summarise(spend = round(mean(totalspend), 2)) %>% 
  spread(Product, spend)
  
```












